Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-02-02 src/ms.tex
JVS comments

\citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}. However, \citet{Curtis2020} demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation. Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, further invalidating the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories.

@@

One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}. However, this scenario is considered implausible as the change in moment of inertia for the Sun over its lifetime is not enough to produce the 5--7 day gap between the Sun and the observed long-period pile-up.

>>>>

\citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}. However, \citet{Curtis2020} demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation. Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, which is in tension with the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories (for sufficiently large volumes).

@@

%One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}. However, this scenario is considered implausible as the change in moment of inertia for the Sun over its lifetime is not enough to produce the 5--7 day gap between the Sun and the observed long-period pile-up.
